Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-05-21 paper_jasa.tex
tweaking intro, and then will rearrange paper to get it into shape for submission

% Article top matter
\title{Validation of Visual Statistical Inference, with Application to Linear Models}
\author{Mahbubul Majumder, Heike Hofmann, Dianne Cook\\
        Department of Statistics, Iowa State University}
%\date{\today}  %\today is replaced with the current date
\maketitle

@@

Statistical graphics plays a vital role in discovery. Finding the unexpected in data, structure that was not previously anticipated or covered by the original hypotheses is one of the strengths of visual analysis. This was an area pioneered by \cite{tukey:eda} and is particularly pertinent in today's data-rich world where data mining has become an important activity. Graphics are also used in many places where numerical summaries simply do not suffice: model checking, diagnosis, and in the communication of findings.

The foundation of statistical graphics is well-established. Work by \cite{cleveland:1984}, repeated and expanded by \cite{heer:2010} provides a framework for evaluating human perception of statistical graphics.  \cite{wilkinson:1999} has developed a grammar of graphics which has been enhanced by \cite{hadley:2009} that formalizes statistics graphics helping to describes how different plots are related. The availability of software systems enabling beautiful data graphics to be tightly coupled with modeling, such as R \citep{R}.

>>>>

% Article top matter
\title{Validation of Visual Statistical Inference, Applied to Linear Models}
\author{Mahbubul Majumder, Heike Hofmann, Dianne Cook\\
        Department of Statistics, Iowa State University}
%\date{\today}  %\today is replaced with the current date
\maketitle

@@

Statistical graphics nourishes the discovery process in data analysis: revealing unexpected things,  finding structure that was not previously anticipated,  or orthogonally contrasting prevailing hypotheses. It is often associated with exploratory data analysis, which was pioneered by \cite{tukey:eda} and, is particularly pertinent in today's data-rich world where discovery during data mining has become an important activity. Graphics are also used in many places where numerical summaries simply do not suffice: model checking, diagnosis, and in the communication of findings.

Statistical graphics research and development is agreeably established now, and has been cemented by several new developments in recent years. Early studies on evaluating how well statistical plots are perceived and read by the human eye \citep{cleveland:1984}, have been repeated and expanded \citep{heer:2010} with findings supporting the original results. This body of work provides a framework for evaluating new statistical graphics.  In a complementary direction, new research on formalizing statistical graphics with language characteristics makes it easier to abstractly define, compare and contrast data plots. \cite{wilkinson:1999} developed a grammar of graphics that is enhanced by \cite{hadley:2009}. These methods provide a mechanism to abstract the way data is mapped to graphical form. Finally, technology advances make it simple and easy for everyone to draw plots of data, and particularly the existence of software systems, such as R \citep{R}, enable making beautiful data graphics that can be tightly coupled with statistical modeling.
